---
title: "Project 872"
author: "Rachael Grainger and Matthew Vining"
date: "2022-11-19"
output: html_document
---

```{r setup, include=FALSE}
#Set Up Document
knitr::opts_chunk$set(echo = TRUE)

#GITHUB LINK: https://github.com/rachael-png/vininggrainger872
```

### Environmental Data Analysis of Hog Farm and Solar Farm Concentration in North Carolina
## By: Rachael Grainger and Matthew Vining
# Fall 2023

# Project Central Question:
How do communities in North Carolina experience varying levels of environmental risks and benefits, with a particular lens on environmental justice communities?

# Potential Hypothesis:
As the birthplace of the environmental justice movement, North Carolina is no stranger to environmental injustices. We expect that today Hog farms will still be correlated/concentrated in communities of higher social vulnerability (i.e., lower SES, predominantly nonwhite) across North Carolina. We also predict that energy systems, such as solar farms, will be concentrated in communities of lower social vulnerability (i.e., higher SES, predominately white) areas in North Carolina.

```{r include=FALSE}
#Load Useful Packages for Project
getwd()
library(dplyr)
library(plyr)
library(lubridate)
library(tidyverse)
library(lubridate)
library(sf)
library(mapview)
library(RColorBrewer)
library(readxl)
#install.packages("viridis")
library(viridis)

#Read in Data

Animal_Operations <- read.csv("Project folder/Raw Data/Animal_Operation_Permits_.csv")

allplant_data <- read_excel("Project folder/Raw Data/allplant_data.xlsx")

solar_data <- read_excel("Project folder/Raw Data/solar_data.xlsx")

#Refine and Combine Data

#filter solar plant data by county and select useful columns
solar_plants_NC <- left_join(solar_data, allplant_data)
solar_plants_NC <- solar_plants_NC %>%
  select(`Utility ID`:Technology, `Nameplate Capacity (MW)`:`Winter Capacity (MW)`, `Street Address`:Longitude) %>%
  group_by(County)


#filter animal farm data for hog farms only and group by county
library(dplyr)
hog_farms <- Animal_Operations %>%
  filter(Type == "Swine State COC") %>%
  group_by(County)


```

## Social Vulnerability Index

CDC's Social Vulnerability Index data will be used to compare areas in North Carolina. 
Social Vulnerability Data: - <https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/SVI_documentation_2018.html> - <https://www.atsdr.cdc.gov/placeandhealth/svi/documentation/pdf/SVI2018Documentation-H.pdf>

Brief notes: - "E\_" are estimates; "M\_" are margins of error - "EP\_ are estimates, in percentages - "SPL_THEME1" is sum of series - "RPL_THEME1" is percentile ranking

We'll focus on just a few variables: Estimates of people in poverty ("E_POV") and of minority population ("E_MINRTY"), keeping the location attributes as well.

```{r join.attributes.to.spatial features, out.width = '100%'}

#Read the 2018 SVI county-level dataset for NC
svi2018_nc_raw <- read.csv(
  'https://svi.cdc.gov/Documents/Data/2018_SVI_Data/CSV/States_Counties/NorthCarolina_COUNTY.csv',
  colClasses = c('FIPS' = 'factor')) %>% 
  select(COUNTY, FIPS, LOCATION, E_TOTPOP, E_POV, E_MINRTY)

#Check structure
str(svi2018_nc_raw)

##Create counties
counties_sf<- st_read('Project folder/Spatial/cb_2018_us_county_20m.shp') %>% 
  filter(STATEFP == 37) #Filter for just NC Counties

#Join the SVI attributes to the county spatial features
counties_sf_join <-  merge(x = counties_sf,
                           y = svi2018_nc_raw, 
                           by.x = "GEOID", 
                           by.y = "FIPS" )

#Tidyverse version of the join
counties_sf_join <- counties_sf %>% 
  left_join(svi2018_nc_raw, by = c("GEOID" = "FIPS") )



#View with mapview
library(leaflet)
counties_sf_join_converted <- st_transform(counties_sf_join, c=4326)

##get percents of poverty
counties_percent <- counties_sf_join_converted %>% group_by(LOCATION) %>% mutate(
                                  PovertyPerc=E_POV/E_TOTPOP, MinorityPerc=E_MINRTY/E_TOTPOP)



##Create Maps of Poverty and Minority Indicators per County



## PERCENT IN POVERTY MAP BY COUNTY
qpal = colorNumeric(palette = "plasma", counties_sf_join_converted$PovertyPerc, reverse = TRUE)

countymap_poverty <- leaflet(counties_percent) %>%
  addPolygons(stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal(PovertyPerc),weight = 1) %>%
  addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=10))  %>%
  addLegend("bottomright", values=~PovertyPerc*100,pal=qpal,title="Percent in Poverty", labFormat = labelFormat(suffix = "%"), opacity = 1)

countymap_poverty

## PERCENT MINORITY MAP BY COUNTY
qpal = colorNumeric( palette = "plasma", domain = counties_sf_join_converted$MinorityPerc, reverse = TRUE)

countymap_minority <- leaflet(counties_percent) %>%
  addPolygons(stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal(MinorityPerc),weight = 1) %>%
  addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=10)) %>%
  addLegend("bottomright", values=~MinorityPerc*100,pal=qpal,title="Percent Minority", labFormat = labelFormat(suffix = "%"), opacity = 1)

countymap_minority


```



```{r hog farm map, out.width = '100%'}

##Convert hog farm data
       
hog_farms.sf  <- hog_farms %>% 
  st_as_sf(coords = c('Longitude','Latitude'),
           crs=4326)

hog_farms.sf2 <- hog_farms.sf %>% drop_na(X)

##Map Hog Farms

##Add Counties Hog Farms
qpal2 = colorNumeric( palette = "plasma", domain = counties_sf_join_converted$MinorityPerc, reverse = TRUE)
qpal3 = colorNumeric( palette = "plasma", domain = counties_sf_join_converted$PovertyPerc, reverse = TRUE)
qpal <- colorNumeric(palette = "plasma", domain = counties_sf_join_converted$PovertyPerc, reverse = TRUE)
paltype <- colorFactor(
   palette = "#ffe9e3",
   domain = hog_farms.sf2$Type)

# Minority Level and Hog Farms
hog_farm_map1 <- leaflet() %>% 
  addPolygons(data =counties_percent, stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal(MinorityPerc),weight = 1) %>%
   addLegend("topright", data = hog_farms.sf2, pal=paltype, values=~Type,title="Farm Type") %>%
    addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=14))  %>%
    addLabelOnlyMarkers(data = hog_farms.sf2, label = ~Facility) %>%
  addLegend("bottomright", data=counties_percent, values=~MinorityPerc*100,pal=qpal2,title="Percent Minority", labFormat = labelFormat(suffix = "%"), opacity = 1) %>%
  addCircleMarkers(data = hog_farms.sf2, radius = 3, color = "black", stroke = TRUE, weight = 1, fillOpacity = 0.8, fillColor = "#ffe9e3")

hog_farm_map1


## Poverty Level and Hog Farms

hog_farm_map_pov <- leaflet() %>% 
  addPolygons(data =counties_percent, stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal3(PovertyPerc),weight = 1) %>%
   addLegend("topright", data = hog_farms.sf2, pal=paltype, values=~Type,title="Farm Type") %>%
    addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=14))  %>%
    addLabelOnlyMarkers(data = hog_farms.sf2, label = ~Facility) %>%
  addLegend("bottomright", data=counties_percent, values=~PovertyPerc*100,pal=qpal3,title="Percent in Poverty", labFormat = labelFormat(suffix = "%"), opacity = 1) %>%
  addCircleMarkers(data = hog_farms.sf2, radius = 3, color = "black", stroke = TRUE, weight = 1, fillOpacity = 0.8, fillColor = "#ffe9e3")

hog_farm_map_pov


```

```{r solar farm maps, out.width = '100%'}

##Convert data to spatial
       
solar_plants_NC.sf  <- solar_plants_NC %>% 
  st_as_sf(coords = c('Longitude','Latitude'),
           crs=4326)

##Map Solar Farms with County Demographics

##Solar + Minority WIHTOUT HOG FARMS

paltype_solar <- colorFactor(
   palette = "#c9ffc7",
   domain = solar_plants_NC.sf$Technology)

solar_map1 <- leaflet() %>% 
  addLegend(data = solar_plants_NC.sf, values=~Technology,pal=paltype_solar,title="Solar Farm Type" ) %>%
  addPolygons(data =counties_percent, stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal(MinorityPerc),weight = 1) %>%
    addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=14))  %>%
addLegend("bottomright", data=counties_percent, values=~MinorityPerc*100,pal=qpal2,title="Percent Minority", labFormat = labelFormat(suffix = "%"), opacity = 1)  %>%
  addCircleMarkers(data = solar_plants_NC.sf, radius = 3, color = "black", stroke = TRUE, weight = 1, fillOpacity = 0.8, fillColor = "#c9ffc7")

solar_map1

##Solar + POVERTY WIHTOUT HOG FARMS

solar_map_pov <- leaflet() %>% 
  addLegend(data = solar_plants_NC.sf, values=~Technology,pal=paltype_solar,title="Solar Farm Type" ) %>%
  addPolygons(data =counties_percent, stroke = TRUE,opacity = 1,fillOpacity = 0.5, smoothFactor = 0.5,
              color="black",fillColor = ~qpal3(PovertyPerc),weight = 1) %>%
    addProviderTiles("Esri.WorldGrayCanvas",
        options = providerTileOptions(minZoom=6, maxZoom=14))  %>%
addLegend("bottomright", data=counties_percent, values=~PovertyPerc*100,pal=qpal3,title="Percent in Poverty", labFormat = labelFormat(suffix = "%"), opacity = 1)  %>%
  addCircleMarkers(data = solar_plants_NC.sf, radius = 3, color = "black", stroke = TRUE, weight = 1, fillOpacity = 0.8, fillColor = "#c9ffc7")

solar_map_pov


```

```{r correlations}

#Prepare hog farm data for correlation testing

Hog_SVI_merge <- hog_farms %>%
  left_join(counties_percent, by = c("County" = "COUNTY") )

Hog_SVI_merge2 <- hog_farms %>%
  group_by(County)

Hog_SVI_merge2  <- aggregate(hog_farms$Type, by=list(hog_farms$County), FUN=length)
colnames(Hog_SVI_merge2) <- c('County','Farms')

#Merge again with SVI data
Hog_SVI_cor <- Hog_SVI_merge2 %>%
  left_join(counties_percent, by = c("County" = "COUNTY"))

#Cor test hog farms

cor.test(Hog_SVI_cor$PovertyPerc, Hog_SVI_cor$Farms)
cor.test(Hog_SVI_cor$MinorityPerc, Hog_SVI_cor$Farms)
summary(aov(Hog_SVI_cor$MinorityPerc ~ Hog_SVI_cor$Farms))
summary(aov(Hog_SVI_cor$PovertyPerc ~ Hog_SVI_cor$Farms))

## Map Hog Farms

# Hog Farms and Poverty Level
plothogpv <- lm(Farms~PovertyPerc,data=Hog_SVI_cor) 
	summary(plothogpv)

plothogpov <-	ggplot(Hog_SVI_cor, aes(x=PovertyPerc, y=Farms)) +
              geom_point() +
              geom_smooth(method=lm, se=FALSE) +
              xlab("Percent Poverty Rate by County") + 
              ylab("Number of Hog Farms") +
              ggtitle("Correlation of Hog Farms & Poverty Levels by County in North Carolina")
             

hogplot1 <- plothogpov + theme(
plot.title = element_text(color="#ba2b04", size=11, face="bold", family="mono"),
axis.title.x = element_text(color="#292929", size=10, face="bold",  family="mono"),
axis.title.y = element_text(color="#292929", size=10, face="bold", family="mono")
)
	
hogplot1

## Hog Farms and Minority Level

plothogmin <-	ggplot(Hog_SVI_cor, aes(x=MinorityPerc, y=Farms)) +
              geom_point() +
              geom_smooth(method=lm, se=FALSE)

plothogmin <- lm(Farms~MinorityPerc,data=Hog_SVI_cor) 
	summary(plothogpv)

plothogmin <-	ggplot(Hog_SVI_cor, aes(x=MinorityPerc, y=Farms)) +
              geom_point() +
              geom_smooth(method=lm, se=FALSE) +
              xlab("Percent Minority Concentration by County") + 
              ylab("Number of Hog Farms") +
              ggtitle("Correlation of Hog Farms & Minority Concentration by County in NC")
             

hogplot2 <- plothogmin + theme(
plot.title = element_text(color="#ba2b04", size=11, face="bold", family="mono"),
axis.title.x = element_text(color="#292929", size=10, face="bold",  family="mono"),
axis.title.y = element_text(color="#292929", size=10, face="bold", family="mono")
)
	
hogplot2
	
#Prepare data for solar farm correlation testing

Solar_SVI_merge <- solar_plants_NC %>%
  left_join(counties_percent, by = c("County" = "COUNTY"))

#Correlation of Solar Farm Capacity and Poverty Level

cor.test(Solar_SVI_merge$PovertyPerc, Solar_SVI_merge$`Nameplate Capacity (MW)`) 

Solar_SVI_merge %>%
  group_by(County) %>%
  summarize(cor=cor(PovertyPerc, `Nameplate Capacity (MW)`))

#Correlation of Solar Farm Capacity and Minority Level
cor.test(Solar_SVI_merge$MinorityPerc, Solar_SVI_merge$`Nameplate Capacity (MW)`) 

Solar_SVI_merge %>%
  group_by(County) %>%
  summarize(cor=cor(MinorityPerc, `Nameplate Capacity (MW)`))

## Solar Plots

## Minority Concentration & Solar Farms

solarmin <- lm(`Nameplate Capacity (MW)`~MinorityPerc,data=Solar_SVI_merge) 
	summary(solarmin)

solarmin_plot <-	ggplot(Solar_SVI_merge, aes(x=MinorityPerc, y=`Nameplate Capacity (MW)`)) +
              geom_point() +
              geom_smooth(method=lm, se=FALSE) +
              xlab("Percent Minority Concentration by County") + 
              ylab("Number of Solar Farms") +
              ggtitle("Correlation of Solar Farms & Minority Concentration by County in NC")
             

solarmin_plot2 <- solarmin_plot + theme(
plot.title = element_text(color="#3cba6c", size=11, face="bold", family="mono"),
axis.title.x = element_text(color="#292929", size=10, face="bold",  family="mono"),
axis.title.y = element_text(color="#292929", size=10, face="bold", family="mono")
)
	
solarmin_plot2

## Poverty & Solar Farms

solarpov <- lm(`Nameplate Capacity (MW)`~PovertyPerc,data=Solar_SVI_merge) 
	summary(solarpov)

solarpov_plot <-	ggplot(Solar_SVI_merge, aes(x=PovertyPerc, y=`Nameplate Capacity (MW)`)) +
              geom_point() +
              geom_smooth(method=lm, se=FALSE) +
              xlab("Percent Poverty by County") + 
              ylab("Number of Solar Farms") +
              ggtitle("Correlation of Solar Farms & Poverty by County in NC")
             

solarpov_plot2 <- solarpov_plot + theme(
plot.title = element_text(color="#3cba6c", size=11, face="bold", family="mono"),
axis.title.x = element_text(color="#292929", size=10, face="bold",  family="mono"),
axis.title.y = element_text(color="#292929", size=10, face="bold", family="mono")
)
	
solarpov_plot2

```
### Summary & Conclusion

While our spatial analysis clearly indicates a concentration of hog farms in counties experiencing high poverty rates, we can decided to run a correlation and regression analysis to test the statistical significance of this concentration. We found that there is a significant presence of hog farms concentrated in counties with higher poverty levels  (p-value = 0.00472). However, the R-squared value is fairly small (0.1165), which indicates there are likely other variables at play that are associated with where hog farms are concentrated in North Carolina besides just poverty levels of the county. Additionally, there is a slight correlation between hog farm concentration and higher minority concentrations, but the p-value is not significant (0.21). 

When analyzing the relationship between solar farms, or their capacity (MW), and poverty and minority levels, spatial analysis revealed that there was not a signficicant relationship. To further explore this relationship, we performed correlation and regression analysis on nameplate capacity of solar farms vs the poverty level and minority level, grouped at the county level to control for differences amongst counties and county borders. The resulting R-squared value for the realtionship with percent minority was very low at 0.004206, meaning there wasn't indication of a strong relationship between electricity capacity and % minority of a county. Further, the p-value was not significant (0.089758), meaning there was not enough evidence to reject a null hypothesis of no relationship. Similarly the relationship between percent poverty and nameplate capapcity was weak, with an R-squared value of 0.004206. The p-value for this relationship was insignificant at .0808.

Overall, we conclude that there is a significant relationship between hog farm placement and the percentage of people in poverty living in a given county. The higher the percentage of impoverished people live in a given county, the greater chance there will be more hog farms located in that county. The relationships explored between hog farms and percent minority, solar farms and percent poverty, and solar farms and percent minority all yielded insignificant statistical analyses. The team recommends adding additional factors outside of poverty and minority measures to understand the underlying factors contributing to placement of hog farms and solar farms. If a broader analysis is done, incorporating more variables, a stronger relationship may be revealed. 

### TO DOUBLE CHECK
Metadata file(s) contains information about data source,
data structure, units of measure, and repository compilers

Flow between text and visualizations is cohesive

Major findings are summarized
Conclusions relate back to the original research context

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


